Jane, John … Leslie? A Historical Method for Algorithmic Gender Detection

### 2014-09-19 article.Rmd
Minor change to abstract

This paper addresses two trends in the humanities: the rise of gender as an analytical tool and the proliferation of large, machine-readable datasets. It outlines a new method for algorithmically inferring the gender of a name from historical data in order to take into account how naming practices change over time and space. In contrast to existing methods of predicting gender, this method provides a higher level of accuracy and a precise statement of the likelihood that a name is a given gender. The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/). It goes on to apply the method to a case study in which we study gender and gatekeeping in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.

>>>>

This paper addresses two trends in the humanities: the rise of gender as an analytical tool and the proliferation of large, machine-readable datasets. It outlines a new method for algorithmically inferring the gender of a name from historical data in order to take into account how naming practices change over time and space. In contrast to existing methods of predicting gender, this method provides a higher level of accuracy and a precise statement of the likelihood that a name was a given gender. The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/). It goes on to apply the method to a case study in which we study gender and gatekeeping in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
